Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome

### 2023-05-31 paper/paper.qmd
Create table of sample counts

---
title: "Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome"
bibliography: references.bib
fontsize: 11pt
date: last-modified
date-format: medium
format:
  manuscript-pdf:
    mainfont: Helvetica
    keep-tex: true
    linenumbers: true
    doublespacing: true
    lof: false
  html:
    embed-resources: true
runningtitle: Predicting CDI Severity from OTUs
runningauthor: Sovacool
papertype: Preprint
author:
  - name: Kelly L. Sovacool
    orcid: 0000-0003-3283-829X
    affiliations:
      - ref: bioinf
  - name: Sarah E. Tomkovich
    affiliations:
      - ref: micro
  - name: Megan L. Coden
    affiliations:
      - ref: mcdb
  - name: Vincent B. Young
    affiliations:
      - ref: micro
      - ref: intmed
  - name: Krishna Rao
    affiliations:
      - ref: intmed
  - name: Patrick D. Schloss
    orcid: 0000-0002-6935-4275
    email: pschloss@umich.edu
    corresponding: true
    affiliations:
      - ref: micro
      - ref: ccmb
affiliations:
  - id: bioinf
    name: Department of Computational Medicine & Bioinformatics, University of Michigan
  - id: micro
    name: Department of Microbiology & Immunology, University of Michigan
  - id: mcdb
    name: Department of Molecular, Cellular, and Developmental Biology, University of Michigan
  - id: intmed
    name: Division of Infectious Diseases, Department of Internal Medicine, University of Michigan
  - id: ccmb
    name: Center for Computational Medicine and Bioinformatics, University of Michigan
#abstract: TODO
#importance: TODO
keywords: _C. difficile_ infection, supervised machine learning, gut microbiome, amplicion sequencing
---
```{r setup, eval=TRUE, echo=FALSE, cache=FALSE}
schtools::set_knitr_opts()
knitr::opts_chunk$set(dpi = 600)
```
```{r load_data, message = FALSE}
library(here)
library(knitr)
library(tidyverse)

load(here("results", "stats.RData"))
```

@@

We first set out to train the best models possible for each severity definition.
Not all samples have labels available for all four severity definitions due to
missing data for some patient lab values and incomplete chart review (TODO table).

@@

@@

table of NNS at 95th pct of risk for each outcome x dataset (@tbl-risk).

```{r risk_95th_pct_decision_thresholds, message=FALSE}
#| label: tbl-risk
#| tbl-cap: "**Predictive model performance at 95th percentile of risk.** The confusion matrix was computed at the 95th percentile of risk for each severity prediction model, which corresponds to 5% of cases predicted to be severe."
read_csv(here('results','decision_thresholds.csv')) %>% kable()
```

@@

decision curve analysis.

>>>>

---
title: "Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome"
bibliography: references.bib
fontsize: 11pt
date: last-modified
date-format: medium
format:
  manuscript-pdf:
    mainfont: Helvetica
    keep-tex: true
    linenumbers: true
    doublespacing: true
    lof: false
  html:
    embed-resources: true
  gfm:
    prefer-html: true
runningtitle: Predicting CDI Severity from OTUs
runningauthor: Sovacool
papertype: Preprint
author:
  - name: Kelly L. Sovacool
    orcid: 0000-0003-3283-829X
    affiliations:
      - ref: bioinf
  - name: Sarah E. Tomkovich
    affiliations:
      - ref: micro
  - name: Megan L. Coden
    affiliations:
      - ref: mcdb
  - name: Vincent B. Young
    affiliations:
      - ref: micro
      - ref: intmed
  - name: Krishna Rao
    affiliations:
      - ref: intmed
  - name: Patrick D. Schloss
    orcid: 0000-0002-6935-4275
    email: pschloss@umich.edu
    corresponding: true
    affiliations:
      - ref: micro
      - ref: ccmb
affiliations:
  - id: bioinf
    name: Department of Computational Medicine & Bioinformatics, University of Michigan
  - id: micro
    name: Department of Microbiology & Immunology, University of Michigan
  - id: mcdb
    name: Department of Molecular, Cellular, and Developmental Biology, University of Michigan
  - id: intmed
    name: Division of Infectious Diseases, Department of Internal Medicine, University of Michigan
  - id: ccmb
    name: Center for Computational Medicine and Bioinformatics, University of Michigan
#abstract: TODO
#importance: TODO
keywords: _C. difficile_ infection, supervised machine learning, gut microbiome, amplicion sequencing
---
```{r setup, eval=TRUE, echo=FALSE, cache=FALSE}
schtools::set_knitr_opts()
knitr::opts_chunk$set(dpi = 600)
```
```{r load_data, message = FALSE}
library(here)
library(kableExtra)
library(knitr)
library(tidyverse)

load(here("results", "stats.RData"))
table_font_size <- 11
```

@@

We first set out to train the best models possible for each severity definition.
Not all samples have labels available for all four severity definitions due to
missing data for some patient lab values and incomplete chart review (@tbl-counts).

@@

```{r sample_counts}
#| label: tbl-counts
#| tbl-pos: h
#| tbl-cap: "Dataset sample counts"
#| tbl-subcap: 
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"
#| layout-ncol: 2
format_table <- function(tab) {
  tab %>%
    kable(col.names = c(' ', 'IDSA', 'Attrib', 'All-cause', "Pragmatic"),
          format.args = list(big.mark = ',')) %>% 
    kable_styling(font_size = table_font_size)
}
read_csv(here('results','count_table_full.csv')) %>% 
  format_table()
read_csv(here('results','count_table_int.csv')) %>% 
  format_table()
```

@@

it's not enough for models to perform well to deploy it in a clinical setting;
benefit over current practices must be shown.
table of NNS at 95th pct of risk for each outcome x dataset (@tbl-risk).

@@

```{r risk_95th_pct_decision_thresholds, message=FALSE}
#| label: tbl-risk
#| tbl-pos: h
#| tbl-cap: "**Predictive model performance at 95th percentile of risk.** 
#|   The confusion matrix was computed at the 95th percentile of risk for each
#|    severity prediction model, which corresponds to 5% of cases predicted to 
#|    be severe. The number needed to screen (NNS) to identify one true positive
#|    is the reciprocal of precision.
#|    "
#| tbl-subcap: 
#|   - "Full datasets"
#|   - "Intersection of samples with all labels available"
risk_thresh <- read_csv(here('results','decision_thresholds.csv'))
risk_thresh %>% 
  filter(Dataset == 'Full') %>% 
  select(-Dataset) %>% 
  kable() %>% 
  kable_styling(font_size = table_font_size)
risk_thresh %>% 
  filter(Dataset == 'Intersection') %>% 
  select(-Dataset) %>% 
  kable() %>% 
  kable_styling(font_size = table_font_size)
```
